ALIGNMENT OF LIGHT AND MASS IN LENSING GALAXIES

### 2013-05-27 shapes/thesis.tex

\usepackage[margin=0.5in]{geometry}
\usepackage{multirow}
\usepackage{lscape}

\usepackage{natbib}
\usepackage{natbibmnfix}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{mathrsfs}
\usepackage{subfigure}

@@

%%%%%%%%%%% Still to change: Dominiks PhD!
\title{ALIGNMENT OF LIGHT AND MASS IN LENSING GALAXIES}
\author{Claudio Bruderer}

@@

\begin{abstract}
\noindent Something
\end{abstract}

%\tableofcontents

\section{Introduction}

\subsection{Context/Background/Motivation}
\begin{itemize}
\item Why interesting (lensing can constrain mass distribution (enclosed mass constraint independent from model with only few percent uncertainty e.g. Kochanek 1991?, Wambsganss \& Paczyński 1994?), insights into structure formation)
\item What is to be expected from theory (dm, alternative gravity)
\end{itemize}

\subsection{Anything done before?}
\textbf{\cite{1997ApJ...482..604K}}: Look at 17 lenses at redshifts ranging between 0.1 and $\sim1$. Sample consists of mainly passively evolving elliptical galaxies and a few spirals. Claim that generally projected light and dark matter distributions are strongly correlated and are not far out of alignment ($\lesssim10^{\circ}$), except in cases where there are strong external tidal perturbations (although misalignments are not necessary in these cases). 'Dark matter halos can be significantly flatter than the light (e.g., Sackett et al. 1994; Buote \& Canizares 1994, 1996; also see the review by Sackett 1996)'. To reconstruct the mass they fit singular isothermal ellipsoid (SIE) lens models, which immediately give them the shape. Another model they use is a singular isothermal sphere (SIS) and some external tidal perturbation (shear). As already stated in \cite{1998ApJ...509..561K}, these models sometimes give a bad fit. To obtain a good fit, an another independent source of external shear is needed (external tidal perturbation). They also state that it is known that the lenses Q0957 + 561 (Young et al. 1980, Grogin \& Narayan 1996), PG 1115 + 080 (Keeton \& Kochanek 1997, Kundic et al. 1997a, Schechter et al. 1997, Tonry 1998), B1422 + 231 (Hogg \& Blandford 1994, Kormann et al. 1994, Kundic et al 1997b, Tonry 1998) are subject to significant tidal perturbations by a nearby cluster, group or galaxy (also line-of-sight structure) (--> All lying in a environment? Ask Prasenjit).

\textbf{\cite{1998ApJ...509..561K}}: They used SIEs for the reconstruction of 6 lenses. They find that for many lenses a model with two independent shears (e.g. SIE + 1 external shear) gives a much better fit. Additionally, the fits are much better than just trying to fit different radial mass distributions with the same number of degrees of freedom. Numerical simulations generally produce more elliptical and more triaxial dark matter halos than luminous galaxies (e.g. Warren et al. 1992; Dubinski 1992, 1994). They can however only compare 3 galaxies with eachother.

\textbf{\cite{1998PhDT.........6K}}: Not sure whether everything was already covered in previous papers or not. He modelled observed lenses to determine their shapes and finds that the dark matter halos are generally aligned with the luminous part. Exceptions are however lenses with known tidal disruptions, where there can be a misalignment. He concludes that the halos were modified by interactions with the galaxies. He however does not state in the abstract how he modelled the lenses.

\textbf{\cite{2000ApJ...538L.113N}}:

\textbf{\cite{2002sgdh.conf...62K}}:

\textbf{\cite{2005MNRAS.360.1333W}}:

\textbf{\cite{2008MNRAS.383..857F}}:

\textbf{\cite{2011AstRv...6f..43B}}:

\textbf{\cite{2012ApJ...761..170G}}:

\textbf{\cite{2000ApJ...536..584L}}:

\subsection{New here}
\begin{itemize}
\item Statistics
\item Free-form (explain advantages of free-form over parametric modelling (\cite{2004AJ....127.2604S}, \cite{2012MNRAS.425.3077L}, \cite{2008ApJ...681..814C}), cite GLASS (\cite{2012MNRAS.425.3077L})
\item group versus stellar misalignment with dm halo
\end{itemize}

\section{Data}
\begin{itemize}
\item SPS (Ignacio Ferreras, Dominik, etc. (\cite{2011ApJ...740...97L}))
\item Bundle information in tables (especially what input data was used for the lenses)
\end{itemize}

\begin{landscape}
 \begin{center}
  \begin{tabular}{l c c c c c c c c c}
	\hline
	\multirow{2}{*}{Lens} & \multirow{2}{*}{$z_{L}$} & \multicolumn{2}{c}{Group/Cluster centroid} & & \multicolumn{2}{c}{Other galaxy} & \multirow{2}{*}{Galaxy type} & \multirow{2}{*}{Environment} & \multirow{2}{*}{References}\\ \cline{3-4} \cline{6-7}
	 & & $\Delta\alpha$ ('') & $\Delta\delta$ ('') & & $\Delta\alpha$ ('') & $\Delta\delta$ ('') & & & \\ \hline \hline
	Q0047 & 0.48 & $3.6\pm15.6$ & $43.2\pm28.8$ &   & & & Early-type & G(9) & \cite{2011ApJ...726...84W}\\
	Q0142 & 0.49 & & &   & & & Early-type & - & \cite{2000ApJ...536..584L}\\
	MG0414 & 0.96 & & &   & $-0.385$ & $1.457$ & Early-type & 1 & \cite{1993AJ....105....1S}\\
	B0712 & 0.41 & & &   & $47$ & $-91$ & Early-type & 1 & \cite{2002AJ....123..627F}\\
	HS0818 & 0.39 & & &   & & & see below & 1 & aaaaaaaaa\\%\cite{2000A&A...357L..29H}\\
	RXJ0911 & 0.77 & $-12.0$ & $-39.8$ &    & & & Early-type & C & \cite{2001ApJ...555....1M}\\
	BRI0952 & 0.63 & &   & & & & Early-type & - & aaaaaaaaa\\%\cite{2007A&A...465...51E}\\
	Q0957 & 0.36 & $4.8$ & $2.1$ &   & & & Early-type & C & \cite{1998ApJ...504..661C}\\
	LBQS1009 & 0.87 & & &   & & & Early-type & - & aaaaaaaaa\\%\cite{2004A&A...428..741F}\\
	B1030 & 0.60 & & &   & $-9.78$ & $8.54$ & Early-type & 1 & \cite{2000ApJ...536..584L}\\
	HE1104 & 0.73 & & &   & & & Early-type & - & aaaaaaaaa\\%\cite{2004A&A...428..741F}\\
	PG1115 & 0.31 & $-10.8\pm21$ & $-3.6\pm15.6$ &   & & & Early-type & G(13) & \cite{2011ApJ...726...84W}\\
	B1152 & 0.439 & & &   & & & Unknown & - & see below\\
	B1422 & 0.34 & $61.2\pm25.2$ & $-10.8\pm23.4$ &   & & & Early-type & G(17) & \cite{2011ApJ...726...84W}\\
	SBS1520 & 0.72 & & &   & & & Early-type & G(4) & see below\\
	B1600 & 0.41 & $-8.44286$ & $-1.48571$ &   & & & Late-type & G(7) & \cite{2007AJ....134..668A}\\
	B1608 & 0.63 & $-0.697246$ & $7.96956$ &   & & & Both & G(8) & \cite{2006ApJ...642...30F}\\
	MG2016 & 1.01 & & &   & & & Early-type & C(69) & \cite{2003MNRAS.344..337T}\\
	B2045 & 0.87 & & &   & & & Early-type & - & \cite{1999AJ....117..658F}\\
	HE2149 & 0.60 & $-36.6$ & $14.2$ &   & & & Early-type & G & \cite{2006ApJ...646...85W}\\
	Q2237 & 0.04 & & &   & & & Late-type & - & see below\\ 
	\hline
  \end{tabular}
 \end{center}
\end{landscape}

The pixelated stellar mass maps we use here was reconstructed by \cite{leier11phd}. For each pixel the surface brightness was compared with stellar mass-to-light ratios, which were determined by populations synthesis model. Details for this method can be found in \cite{leier11phd}, \cite{2005ApJ...623L...5F}, \cite{2008MNRAS.383..857F}.

Here some comments about the environments of the lenses in the sample.

\textbf{Q0047-2808}: \cite{2011ApJ...726...84W} find a group of 9 members, spectroscopically confirmed. According to \cite{1996MNRAS.278..139W} it is a luminous early-type galaxy.

\textbf{Q0142-100}: The required external shear is comparable to the cosmic LSS-contribution, and it also seems to be a passively evolving early-type galaxy according to \cite{2000ApJ...536..584L} and confirmed by \cite{2007A&A...465...51E}.

\textbf{MG0414+0534}: According to \cite{1999AJ....117.2034T} the lens seems to be a passively evolving early-type galaxy. \cite{1993AJ....105....1S} find a luminous satellite galaxy north-west of the lens. The position data of this object was taken from CASTLES.

\textbf{B0712+472}: According to \cite{2002AJ....123..627F} potentially only one other galaxy 102'' to the south-east at a similar redshift. There is however a group of 10 galaxies at a lower redshift. According to \cite{1998AJ....115..377F} the spectrum is typical for an early-type galaxy.

\textbf{HS0818+1227}: According to \cite{2000A&A...357L..29H} one galaxy, which seems to be at the same redshift, can be found near the lens. In roughly the same direction, two other brighter galaxies can be found a bit farther away. However, no further inquiries were made whether they all belong to the same group or not. There are also a couple of fainter ones, which could potentially be group members. \underline{TO DO}: Get relative position of nearest galaxy! More or less everybody lists HS0818 as an early-type, however noone states a source for their belief.

\textbf{RXJ0911+0551}: According to \cite{2001ApJ...555....1M} the lens lies on the outskirts of a cluster environment. The cluster seems to be rather complex and not spherical, and it possibly is not dynamically relaxed. The X-ray emission analysis yields a temperature of 2.3 keV. There is also a satellite galaxy to the north-west direction (\cite{2000ApJ...544L..35K}). The lens is a almost circular early-type galaxy (\cite{2012A&A...538A..99S}). Time delay: $\Delta t_{BA}=146\pm4$ (\cite{2002ApJ...572L..11H}; CASTLES: BD-delay, here BA-delay).

\textbf{BRI0952-0115}: Lens possibly part of smaller system (of 2 or 3 other galaxies), which is found at a slightly higher redshift, according to \cite{2007A&A...465...51E}. Surely group found in \cite{2006ApJ...641..169M} is at a lower redshift than the lens as the redshift estimate $z=0.41$ was used. \cite{2007A&A...465...51E} however find that the redshift is more likely to be $z=0.632$. In \cite{2000ApJ...536..584L} it is found that required external shear can be explained easily by a cosmic contribution. \cite{2000ApJ...536..584L}, and later on confirmed by \cite{2007A&A...465...51E}, find that it seems to be a typical early-type galaxy. \underline{TO DO}: Why is the such a hugely different redshift estimate in Table 4 of \cite{2006ApJ...641..169M}? Reason: In \cite{2000ApJ...536..584L} this redshift is listed as the lens redshift. --> Correct redshift used in analysis?

\textbf{Q0957+561}: The lens is a cD galaxy lying close to the center of a cluster with a high sprial galaxy-fraction (e.g. \cite{1992MNRAS.254P..27G}, \cite{1994A&A...291..411A}, \cite{1998ApJ...504..661C}). As listed in table 3 of \cite{2000ApJ...542...74K}, the position angles to the center of the cluster of earlier lens reconstructions range between 51.8 deg and 67.8 deg measured north-through-east. Judging 'by hand' from the center of the X-ray emission measured by \cite{1998ApJ...504..661C}, we find $\Delta\alpha=4.8''$ and $\Delta\delta=2.1''$, which corresponds to a position angle of 66.4 deg, consistent with the values listed above. \cite{2012A&A...540A.132S} measure the time delay in the g- and r-band. They find $\Delta t_{BA}=416.5\pm1.0$ in the g-band, $\Delta t_{BA}=420.6\pm1.9$ in the r-band. So, they find a three-day lag between these two bands and the two estimates do not agree at the 2$\sigma$-level. They argue that this effect can be accounted for by the presence of a substructure and chromatic dispersion (light propagation time increases with decreasing wavelength). \underline{TO DO}: Better measure of the cluster center? Time delay?

\textbf{LBQS1009-0252}: \cite{2000ApJ...536..584L} find that it is probably an early-type galaxy. According to \cite{2004A&A...428..741F}, there is no significant galaxy overdensity near the lens.

\textbf{B1030+074}: As noted by \cite{1998MNRAS.300..649X} it appears to have some substructure. \cite{2000ApJ...536..584L} conclude that they are two distinct galaxies. The main one is an red, early-type galaxy. The other component seems to be fitted well by an exponential disk galaxy. They also conclude that no firm statements about the environment can be made. Although there is another galaxy nearby, it is too blue to be an early-type galaxy and thus they cannot estimate its shear contribution. \underline{TO DO}: No error on position of G1 in \cite{2000ApJ...536..584L}?

\textbf{HE1104-1805}: \cite{2000ApJ...536..584L} find that it is probably an early-type galaxy (also \cite{2000A&A...360..853C}). They also seem to find a group environment. \cite{2004A&A...428..741F} however rather think that their photometric redshift measurement indicate that these galaxies are at a higher redshift and rather associated with the source. As it is also listed in \cite{2004A&A...428..741F}, modelling of the lens, parametric and free-form, requires strong external shear. Time delay: $\Delta t_{AB}=162.2^{+6.3}_{-5.9}$ (\cite{2008ApJ...676...80M}; here BA-delay).

\textbf{PG1115+080}: Environments analyzed by \cite{2006ApJ...641..169M} and \cite{2011ApJ...726...84W}. Lying in a small group of 13 members. The lens is an early-type galaxy (\cite{2005ApJ...626...51Y}). \cite{2004ApJ...610..686G} detect X-ray emission from the corresponding group at temperature 0.8 keV. Time delays: \cite{1997ApJ...489...21B} find $\Delta t_{AC}=13.3^{+0.9}_{-1.0}$ and $\Delta t_{BA}=11.7^{+1.5}_{-1.6}$, \cite{2010MNRAS.406.2764T} find however $\Delta t_{AC}=12.0^{+2.4}_{-2.0}$ and $\Delta t_{BA}=4.4^{+3.2}_{-2.4}$ (here actually BA- and CD-delay).

\textbf{B1152+200}: Not much seems to be known about the environment. The lens galaxy type is not known. According to \cite{2000A&A...357..115T} however, its color is consistent with a late-type or an irregular galaxy.

\textbf{B1422+231}: \cite{2006ApJ...641..169M} find totally 16 spectroscopically confirmed member galaxies of the group. Using newer data, \cite{2011ApJ...726...84W} find an additional member. \cite{2004ApJ...610..686G} detect X-ray emission from the corresponding group at temperature 1.0 keV. According to \cite{1996ApJ...462L..53I} the lens is an early-type galaxy. Although there are measured time delays by \cite{2001MNRAS.326.1403P}, they are probably not to trust too much as they deviate quite a bit from theoretical expectations in \cite{2003AJ....126...29R}. An accurate measurement of the time delays probably requires time delay measurements on the scale of hours.

\textbf{SBS1520+530}: There is some uncertainty about the redshift of the galaxy. \cite{2002A&A...391..481B} find that their spectroscopic redshift estimate of the galaxy is consistent with $z=0.71$, which was measured first by \cite{1997A&A...318L..67C}. Both however also measured a system at $z=0.82$. \cite{2002A&A...391..481B} however conclude that it does not seem to be associated with the lens. \cite{2008ApJ...673..778A} find another system at $z=0.76$. They argue that this system is more likely to be the lens, and the foreground system at $z=0.71$ just a local perturber. The data however does not seem to be conclusive. \cite{2008ApJ...673..778A} also find and spectroscopically confirm 3 groups in the line of sight. One is at $z=0.716$, with 5 members and the centroid position: $\alpha=15:21:48.16$, $\delta=52:53:19.5$. Another one at $z=0.758$, with 13 members and $\alpha=12:21:38.22$, $\delta=52:55:23.1$. Another one is at the higher redshift $z=0.818$ with 10 members. Including the lens would change these positions. The lens seems to be an early-type galaxy, possibly with a disc (\cite{2008ApJ...673..778A}). Time delay: $\Delta t_{BA}=125.8\pm2.1$ (\cite{2011A&A...536A..44E}; here also BA-delay). \underline{TO DO}: Dominik has used a lower redshift. Implications for the stellar population synthesis? Redo analysis at a higher redshift.

\textbf{B1600+434}: The lens is a nearly edge-on late-type galaxy (\cite{1997A&A...317L..39J}). No significant X-ray emission was detected from the surrounding group (\cite{2005ApJ...625..633D}). According to \cite{2007AJ....134..668A}, the lens lies in a group with at least seven members, which all seem to be late-type galaxies. The centroid of the group is not luminosity-weighted in the table. Time delay: $\Delta t_{AB}=51\pm2$ (\cite{2000ApJ...544..117B}; here BA-delay). \underline{TO DO}: No error on position estimate in paper? --> What should the error estimate for the centroid be? Add environment image?

\textbf{B1608+656}: The environment and the mass distribution along the line of sight have been analyzed by \cite{2006ApJ...642...30F}. All data points to a galaxy merger, so we could as well talk about 9 group members. A luminosity weighted measure was taken to find the group centroid. No errors are given, but as a comparision the median centroid is listed: $\Delta\alpha = 12.5$, $\Delta\delta = 21.6506$. Along the line of sight, there seem to be four other groups. No significant X-ray emission was detected from the surrounding group (\cite{2005ApJ...625..633D}). According to \cite{2003ApJ...584..100S} the galaxy G1 is an early-type galaxy which disrupted probably a late-type galaxy G2. Time delays: $\Delta t_{AB}=31.5^{+2}_{-1}$, $\Delta t_{CB}=36.0\pm1.5$, $\Delta t_{DB}=77.0^{+2}_{-1}$ (\cite{2002ApJ...581..823F}; here BA-, CA-, DA-delay). \underline{TO DO}: Errors? Add environment image?

>>>>

\usepackage[margin=0.8in]{geometry}
\usepackage{multirow}
\usepackage{lscape}

\usepackage{natbib}
\usepackage{natbibmnfix}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{mathrsfs}

\usepackage{caption}
\usepackage{subcaption}

@@

\title{ALIGNMENT OF LIGHT AND MASS IN LENSING GALAXIES}
\author{Claudio Bruderer}
\date{}

@@

\begin{abstract}
\noindent Understanding the structure and properties of galaxies today is essential to understand how galaxies formed. This poses a larger problem to us than one might imagine. Especially constraining the dark matter distribution in galaxy halos is not trivial. There are only a few methods to directly estimate this distribution. One of them is strong gravitational lensing, an effect caused by all of mass. This phenomenon allows constraining the gravitational potential of lens galaxies and thus the mass distribution. With knowledge of the light distribution it is possible to disentangle the dark matter and baryonic mass components and study their properties.
We use the newly-developped nonparametric framework \textit{GLASS} to model the mass distribution of a total of 15 strong gravitational lensing galaxies, most of which are early-type galaxies, for which we have information on the baryonic mass distribution. We are thus able to separate the mass components. By using a simple shape measure, namely the inverse of the moment of inertia tensor, we can analyze the shapes of the distributions. Analyzing the alignments between the semi-major axes of the distributions, we find that there can be significant misalignments and also cases of alignment between the semi-minor and semi-major axes. Mass thus need not follow light. Furthermore, the ellipticities of the dark matter distributions are always smaller than the baryonic mass distributions. Dark matter halos tend to be more spherical than the light distribution of the galaxy. We also consider the environment of the lenses. We find that lenses in denser environments tend to be more misaligned with either the semi-major or semi-minor axis. Additionally, the semi-major axis of lenses lying in groups or clusters of the dark matter halo might be oriented in direction of the centroid.
\end{abstract}

\tableofcontents

\clearpage
\section{Introduction}\label{sec:introduction}

Gravitational lensing is one of the few phenomena that allow probing the matter distribution directly. It makes no distinction between baryonic and dark matter, both components act on light rays in the same way. Gravitational lensing is therefore somewhat special as it allows to 'see' dark matter, if we are able to separate the two components. Therefore studying these effects allows us to probe dark matter and its properties. We are typically interested in measuring its distribution, how it clusters, its connection to structure formation, and of course its nature. With strong gravitational lensing galaxies we are effectively probing the matter distribution on relatively small scales. Thus, much insight could be gained in how galaxies form and how they are structured.

The main challenges are however estimating first, the baryonic mass distribution, such that the two componantes can be disentangled, and second, modelling the lens, i.e. the mass distribution. We do not elaborate the first problem (see e.g. \cite{leier11phd}) further but skip right ahead to the second problem. There are in general two classes of methods to reconstruct gravitational lenses, parametric and nonparametric methodes. A comparison of these two classes of methods can be found in \cite{2008ApJ...681..814C}. Parametric methods assume a distribution, which depends only on few parameters, such that a best-fitting model can be found. Nonparamatric (or free-form) methods chose a specific base set. Due to the fact that there is no \textit{a priori} reason to prefer certain mass models over others, given that they all satisfy certain prior conditions and reproduce the same gravitational lensing effects, there is no best-fitting model. It is therefore important to explore many models and look at the properties of all the models in the ensemble. We typically need to sample high-dimensional parameter spaces, which describe all allowed models, in a uniform and uncorrelated way to generate a representative sample. This is a computational challenge and certainly not trivial \citep{2008ApJ...679...17C,2012MNRAS.425.3077L}. Parametric methods are usually computationally cheaper, but as the distributions are assumed beforehand, they might bias the result too much.

With these tools, we can start answering the question whether mass follows light. It is a well-known fact that for observations to match e.g. rotation curves and simulations, the radial density profiles of baryonic mass and dark matter differ. As we know that galaxies can be quite triaxial, the next question to ask is whether dark matter and baryonic matter need to be aligned and have similar ellipticities. If this is not the case, then this needs to be explained by mechanism and effects (e.g. environment of the lenses) that produce and sustain these different properties. It is furthermore unclear whether modified gravity theories introduced to explain the 'missing matter' could explain misalignments between mass and light.

There were already attempts to analyze the alignment of light and mass. \cite{1997ApJ...482..604K} started looking at misalignments quite early on. They use Singular Isothermal Spheres (SIS) with an external shear term and Singular Isothermal Ellipsoids (SIE) (e.g. \cite{1994A&A...284..285K}) without external shear to model the mass distribution. Applying this to a sample of 17 lenses, they find a strong correlation between light and dark matter distributions, while there is only a small scatter of less than $10^{\circ}$, if there is no strong external perturbation. They find however no correlation between the axis ratios of the distributions.

Later on, \cite{2006ApJ...649..599K} and \cite{2012ApJ...761..170G} analyzing lens samples of similar sizes and \cite{2009ApJ...690..670T} analyzing a lens sample containing even 70 galaxies also looked for misalignments. They all use SIEs to model the mass distributions. Only \cite{2012ApJ...761..170G} however include an external shear contribution in case more information is available. They neglecht the external shear although \cite{1998ApJ...509..561K} find that SIEs with an independent external shear term fit the data better. They find reasonable alignments, the scatter is however larger than the one they find in \cite{1997ApJ...482..604K}. \cite{2012ApJ...761..170G} find a rms scatter of $25^{\circ}$ for the whole sample. Even after dropping the almost spherical lenses, the rms scatter is still $18^{\circ}$. They attribute this to a significant external shear contribution. \cite{2009ApJ...690..670T} even find a significant correlation between the misalignment and the local resp. global environment. In general, lenses in overdense environment seem to be more likely to be misaligned. \cite{2009ApJ...690..670T} and \cite{2012ApJ...761..170G} also find a correlation with a large scatter between the ellipticities of the light and dark matter distributions.

Contrary to earlier approaches, we choose to use a non-parametric modelling technique. We use a pixel-based (see e.g. \cite{1997MNRAS.292..148S}) lens reconstruction framework called \textit{GLASS} \citep{2012MNRAS.425.3077L}. We apply this successfully on a sample of 15 strong gravitational lensing galaxies using stellar mass data of \cite{leier11phd}. Thus, we can disentangle dark and luminous matter and look at the shapes and alignments of these distributions. We can also look at the ellipticities of the reconstructed lenses and compare them to the ones of the stellar distribution. Furthermore, we discuss the environments of these lenses, as they might affect some properties of gravitational lenses in a significant way \citep{2006ApJ...641..169M,2011ApJ...726...84W}. We think that we have enough statistics to make claims about misalignments between the dark matter halo and the stellar mass distribution, and how the halo is oriented compared to the environment.

In Section \ref{sec:data} we present the lens sample and important properties of each lens. In Section \ref{sec:method} we present the reconstruction method we use and our shape measure. Section \ref{sec:results} consists of the results of this analysis. We show the relations we find between misalignment, ellipticities, and the environment.

\section{Data}\label{sec:data}
\begin{table}[t]
 \centering
 \begin{minipage}{.64\textwidth} %.63
  \begin{tabular}{l c c l | l c c l}
   Lens & $z_{L}$ & $z_{S}$ & Type & Lens & $z_{L}$ & $z_{S}$ & Type\\ \hline \hline
   Q0047 & 0.48 & 3.60 & Quad & PG1115 & 0.31 & 1.72 & Quad\\
   Q0142 & 0.49 & 2.72 & Double & B1152 & 0.44 & 1.02 & Double\\
   MG0414 & 0.96 & 2.64 & Quad & B1422 & 0.34 & 3.62 & Quad\\
   B0712 & 0.41 & 1.34 & Quad & SBS1520 & 0.72 & 1.86 & Double\\
   RXJ0911 & 0.77 & 2.80 & Quad & B1600 & 0.41 & 1.59 & Double\\
   BRI0952 & 0.63 & 4.50 & Double & B1608 & 0.63 & 1.39 & Quad\\
   Q0957 & 0.36 & 1.41 & Double & MG2016 & 1.01 & 3.27 & Quad$^{1}$\\
   LBQS1009 & 0.87 & 2.74 & Double & B2045 & 0.87 & 1.28 & Quad\\
   B1030 & 0.60 & 1.54 & Double & HE2149 & 0.60 & 2.03 & Double\\
   HE1104 & 0.73 & 2.32 & Double & Q2237 & 0.04 & 1.69 & Quad\\
  \end{tabular}
  \captionsetup{width=\linewidth}
  \captionof{table}{List of all the lenses in the sample and their properties. Data is taken from \textit{CASTLES} and \cite{leier11phd}. The angles are measured north-through-east. \newline $^{1}$ Two images merge, thus only three images can be resolved.}
  \label{tab:overview}
 \end{minipage}
\end{table}

\begin{table}[t]
 \centering
 \begin{minipage}{.83\textwidth} %.8
  \begin{tabular}{l c c c c c c c}
   \multirow{2}{*}{Lens} & \multirow{2}{*}{Environment$^{a}$} & \multicolumn{2}{c}{Group/Cluster centroid$^{b}$} & & \multicolumn{2}{c}{Other galaxy} & \multirow{2}{*}{References$^{c}$}\\ \cline{3-4} \cline{6-7}
    & & $\Delta\alpha$ [''] & $\Delta\delta$ [''] & & $\Delta\alpha$ [''] & $\Delta\delta$ ['']\\ \hline \hline
   Q0047 & G(9) & $3.6\pm15.6$ & $43.2\pm28.8$ &   & ... & ... & 1\\
   MG0414 & 2 & ... & ... &   & $-0.4$ & $1.5$ & 2\\
   B0712 & 2 & ... & ... &   & $47$ & $-91$ & 3\\
   RXJ0911 & C & $-12.0$ & $-39.8$ &    & ... & ... & 4\\
   Q0957 & C & ... & ... &   & ... & ... & 5\\
   B1030 & 2 & ... & ... &   & $-9.8$ & $8.5$ & 6\\
   PG1115 & G(13) & $-10.8\pm21.0$ & $-3.6\pm15.6$ &   & ... & ... & 1\\
   B1422 & G(17) & $61.2\pm25.2$ & $-10.8\pm23.4$ &   & ... & ... & 1\\
   SBS1520 & G(5) & ... & ... &   & ... & ... & 7\\
   B1600 & G(17) & $-8.4^{d}$ & $-1.5^{d}$ &   & ... & ... & 8\\
   B1608 & G(8) & $-0.7$ & $8.0$ &   & ... & ... & 9\\
   MG2016 & C(69) & ... & ... &   & ... & ... & 10\\
   HE2149 & G & $-36.6$ & $14.2$ &   & ... & ... & 11\\
  \end{tabular}
  \captionsetup{width=\linewidth}
  \captionof{table}{List of lenses lying in a group or cluster environment, or that have a companion. \newline $^{a}$ G stands for group, C for cluster, and 2 means that the lens has a companion galaxy. If known, the number of member galaxies including the lens is given. \newline $^{b}$ If not stated differently, the centroid positions are luminosity-weighted estimates. \newline $^{c}$ References for the centroid estimates we used. \newline $^{d}$ Only number-weighted estimates are available. \newline 1: \cite{2011ApJ...726...84W}; 2: \cite{1993AJ....105....1S}; 3: \cite{2002AJ....123..627F}; 4: \cite{2001ApJ...555....1M}; 5: \cite{1998ApJ...504..661C}; 6: \cite{2000ApJ...536..584L}; 7: \cite{2008ApJ...673..778A}; 8: \cite{2007AJ....134..668A}; 9: \cite{2003MNRAS.344..337T}; 10: \cite{2006ApJ...646...85W}; 11: \cite{2006ApJ...646...85W}}
  \label{tab:environment}
 \end{minipage}
\end{table}

\begin{table}[t]
 \centering
 \begin{minipage}{.79\textwidth} %.77
  \begin{tabular}{l c c c c}
   Lens & \multicolumn{3}{c}{Time delays [d]} & References\\ \hline \hline
   RXJ0911 & $\Delta t_{BA}=146\pm4$ & ... & ... & 1\\
   Q0957 & $\Delta t_{BA}=416.5\pm1.0^{a}$ & $\Delta t_{BA}=420.6\pm1.9^{b}$ & ... & 2\\
   HE1104 & $\Delta t_{BA}=162.2^{+6.3}_{-5.9}$ & ...  & ...  & 3\\
   PG1115 & $\Delta t_{BA}=12.0^{+2.4}_{-2.0}$ & $\Delta t_{DC}=4.4^{+3.2}_{-2.4}$ & ... & 4\\
   SBS1520 & $\Delta t_{BA}=125.8\pm2.1$ & ... & ... & 5\\
   B1600 & $\Delta t_{BA}=51\pm2$ & ... & ... & 6\\
   B1608 & $\Delta t_{BA}=31.5^{+2}_{-1}$ & $\Delta t_{CA}=36.0\pm1.5$ & $\Delta t_{DA}=77.0^{+2}_{-1}$ & 7\\
   HE2149 & $\Delta t_{BA}=103\pm12$ & ... & ... & 8\\
  \end{tabular}
  \captionsetup{width=\linewidth}
  \captionof{table}{Lenses with measured time delays. $\Delta t_{BA}$ corresponds to the time delay between the images A and B with A leading. \newline $^{a}$ measured in the g-band \newline $^{b}$ measured in the r-band \newline 1: \cite{2002ApJ...572L..11H}; 2: \cite{2012A&A...540A.132S}; 3: \cite{2008ApJ...676...80M}; 4: \cite{2010MNRAS.406.2764T}; 5: \cite{2011A&A...536A..44E}; 6: \cite{2000ApJ...544..117B}; 7: \cite{2002ApJ...581..823F}; 8: \cite{2002A&A...383...71B}}
  \label{tab:timedelays}
 \end{minipage}
\end{table}

The pixelated stellar mass maps we use, were reconstructed by \cite{leier11phd}. For each pixel the surface brightness was compared with stellar mass-to-light ratios, which were determined by stellar population synthesis models. This technique allows estimating the stellar mass in these pixels. Further details on this method can be found in \cite{leier11phd}, \cite{2005ApJ...623L...5F}, \cite{2008MNRAS.383..857F}.

Our sample consists of 20 galaxies. The sample is identical to the one as in \cite{leier11phd} and \cite{2011ApJ...740...97L} except that \textit{HS0818+1227} proved too difficult to reconstruct and therefore was left out. Most of the lens galaxies are early-type galaxies, except \textit{B1030+074}, \textit{B1152+200}, \textit{B1600+434}, \textit{B1608+656}, and \textit{Q2237+030}. All lenses can also be found in the gravitational lens database \textit{CASTLES}. Characteristics and references of each lens in the sample are listed below. Table \ref{tab:overview} contains of all the redshifts and the type of the lenses, Table \ref{tab:environment} lenses with environment information, and Table \ref{tab:timedelays} lenses with measured time delays.

\textit{Q0047-2808} (hereafter \textit{Q0047} or \textit{0047}) is a luminous early-type galaxy \citep{1996MNRAS.278..139W}. \cite{2011ApJ...726...84W} find a group of 9 members all of which they spectroscopically confirm.

\textit{Q0142-100} (hereafter \textit{Q0142} or \textit{0142}) is a passively evolving early-type galaxy according to \citep{2000ApJ...536..584L,2007A&A...465...51E}.

\textit{MG0414+0534} (hereafter \textit{MG0414} or \textit{0414}) is a passively evolving early-type galaxy \citep{1999AJ....117.2034T}. \cite{1993AJ....105....1S} find a very close luminous satellite galaxy north-west of the lens. Due to the proximity of this satellite, we include it as a point mass in our analysis.

\textit{B0712+472} (hereafter \textit{B0712} or \textit{0712}) is an early-type galaxy \citep{1998AJ....115..377F}. Another galaxy, 102'' to the south-east seems to be at a similar redshift \citep{2002AJ....123..627F}. There is also a group of 10 galaxies at a lower redshift.

\textit{RXJ0911+0551} (hereafter \textit{RXJ0911} or \textit{0911}) is an almost circular early-type galaxy \citep{2012A&A...538A..99S}. It has measured time delays \citep{2002ApJ...572L..11H}. It lies on the outskirts of a cluster \citep{2001ApJ...555....1M}. This cluster seems to be rather complex and especially it is not spherical. Possibly it is not dynamically relaxed, although X-ray emission can be detected. Analysis of this emission yields a temperature of 2.3 keV. There is also a satellite galaxy to the north-west direction \citep{2000ApJ...544L..35K}.

\textit{BRI0952-0115} (hereafter \textit{BRI0952} or \textit{0952}) is an early-type galaxy \citep{2000ApJ...536..584L,2007A&A...465...51E}. \cite{2006ApJ...641..169M} find that the lens lies in a poor group of 5 members. However, a redshift of $z=0.41$ was used. According to \cite{2007A&A...465...51E}, the redshift is more likely to be $z=0.632$. They also find a system of 2 or 3 other galaxies at a slightly higher redshift of which the lens could be part of. The required external shear could be explained easily by a cosmic contribution \citep{2000ApJ...536..584L}. Therefore, we treat this lens as an isolated object.

\textit{Q0957+561} (hereafter \textit{Q0957} or \textit{0957}) is a cD galaxy lying close to the center of a cluster with a high sprial galaxy-fraction (e.g. \cite{1992MNRAS.254P..27G,1994A&A...291..411A,1998ApJ...504..661C}). Due to the large image separation, large physical scales are probed. As listed in Table 3 of \cite{2000ApJ...542...74K}, the position angles to the center of the cluster of earlier lens reconstructions range between 51.8 deg and 67.8 deg measured north-through-east. These results are consistent with the center of the X-ray emission from the cluster \citep{1998ApJ...504..661C}. The lens has a measured time delay, e.g. \citep{2012A&A...540A.132S}. They however also find a three-day lag between the g- and r-bands, and the estimates do not agree at the 2$\sigma$-level. They argue that this effect can be accounted for by the presence of a substructure and chromatic dispersion. We find that the results do not change significantly for either estimate and choose therefore the g-band measurement.

\textit{LBQS1009-0252} (hereafter \textit{LBQS1009} or \textit{1009}) is probably an early-type galaxy \citep{2000ApJ...536..584L}. \cite{2004A&A...428..741F} find no significant galaxy overdensity near the lens.
